---
title: Cognitive and Auditory Predictors of Dysarthric Speech Understanding in Older
  Adult Listeners
author: "Lansford, Hirsch, Barrett, & Borrie"
date: "`r Sys.Date()`"
output: 
  github_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This document contains the analysis code and output for the manuscript entitled "Cognitive and Auditory Predictors of Dysarthric Speech Understanding in Older Adult Listeners". These results will also be presented at the Biannual Motor Speech Conference held in San Diego, CA from February 21-24, 2024.

All data processing and analyses was conducted using R Version 4.3.2


```{r, echo = F, warning = F, message = F}

# Loading in needed packages
library(tidyverse) #install.packages("tidyverse")
library(rio) #install.packages("rio")
library(gt) #install.packages("gt")
library(gtsummary) #install.packages("gtsummary")
library(reshape2) #install.packages("reshape2")
library(ggcorrplot) #install.packages("ggcorrplot")
library(patchwork) #install.packages("patchwork")
library(janitor) #install.packages("janitor")
library(fs) #install.packages("fs")
library(furniture) #install.packages("furniture")
library(caret) #install.packages("caret")
library(effectsize) #install.packages("effectsize")
library(parameters) #install.packages("parameters")
library(performance) #install.packages("performance")
library(bayestestR) #install.packages("bayestestR)
library(rstanarm) #install.packages("rstanarm")

# Load in dataset

demo <- rio::import("participant_demo.csv")

data <- rio::import("cleaned_data.csv")

```


```{r, echo = F, warning = F, message = F}

# Functions used in this code

## Function for determining variable importance
var_imp <- function(mod, label) {
  varImp(mod)$importance %>%
    rownames_to_column(var = "Variable") %>%
    mutate(Overall = Overall/max(Overall)) %>%
    rename(!!label := Overall, 
          var = Variable)
}

## Function for model training
training <- function(data, method, int) {
  form = if (isTRUE(int)) {
    d2 = model.matrix(pwc ~ . * ., data = data)[, -1] %>%
      as_tibble() %>%
      janitor::clean_names()
    d2$pwc = data$pwc
  } else {
    d2 = data
    train(
      pwc ~.,
      data = d2,
      preProcess = c("center", "scale"),
      method = "repeatedcv",
      number = 5,
      repeats = 10
    )
  }
}

```


```{r}

# Calculating pretest and posttest intelligibility for the listeners
pl_int <- data %>%
  dplyr::group_by(id, speaker, type) %>%
  dplyr::mutate(pwc = (sum(correct_words)/sum(target_number))*100,
                best_win = ifelse(win_r < win_l, win_r, win_l)) %>%
  dplyr::summarize(mean_pwc = mean(pwc, na.rm = T),
                   across(c(contains("_c"), contains("_u"), contains("win")),
                          ~first(.x)
                          )) %>%
  dplyr::ungroup()

# Subsetting age from demo df

age <- demo %>%
  select(id, age)

# Merging listener age into pl_int

pl_int <- dplyr::left_join(age, pl_int, by = "id")

```



# Descriptives

## Participant Demographics

TBD

```{r}



```

## Cognition

```{r}

int_table <- pl_int %>%
  dplyr::filter(type == "pretest") %>%
  dplyr::select(id, speaker, age, ends_with("_c"), best_win) %>%
  dplyr::select(-fluid_cog_c) %>%
  dplyr::rename(
    `Working Memory` = list_sort_c,
    `Inhibitory Control` = flanker_c,
    `Cognitive Flexibility` = card_sort_c,
    `Receptive Vocabulary` = vocab_c,
    `Processing Speed` = pattern_c,
    `Picture Sequence Memory` = pic_seq_c,
    `Words in Noise Threshold` = best_win
  ) %>%
  group_by(id, speaker) %>% 
  summarize(across(everything(), first)) %>% 
  ungroup() %>% 
  select(-id) %>% 
  tbl_summary(
    by = "speaker",
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no",
    digits = list(everything() ~ c(2))
  ) %>% 
  add_p() %>%
  as_gt() 

int_table

int_table %>%
  gt::gtsave("Tables/means_tabel.html")

```

## Intelligibility

```{r}

# Intelligibiltiy Descriptives Table



```


